### 2021-12-09 documentos/generators.tex
Medio corregido el párrafo de la caracterización. Falta explicación rigurosa

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
%\janote{Este párrafo lo voy a retocar. El enunciado de arriba es el primer enunciado
%del nuevo párrafo. Ojo que hay que introducir la notación $\mcG$.}
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs can be obtained as
compositions the PCEs corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCEs that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
% and the complete vector space (corresponding to the identity channel),
% which we call generators. 
% \cpnote{Aca tenemos un semigrupo y no se si incluir a la identidad como 
% generador. De alguna manera es excepcional. Es problematico incluirla y es 
% problematico no incluirla.}
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
generators and the composition of any two of them yield the PCE channel
corresponding to (b). 
% 
% 
% the composition of any two 2-qubits PCE
% generators 
% in (c), (d), or (e) is equivalent to taking the intersection between them both
% and the three different possible compositions result in PCE channel in (b).
\cpnote{Aca se me hace que hay un cambio brusco de tema. Ver que onda
con Jose Alfredo}
The existence of the PCE generators $\mcG$ provide the program of 
constructing all PCE channels a tool that avoids the usage of a 
high-computational-cost method. In a moment, we will discuss how to find the 
generators $\mcG$ from matrix $A$, given a symmetry that is obeyed both by the generators 
$\mcG$ and matrix $A$ in \Eref{eq:san}\cpnote{Como que para que ahora se
menciona entonces?}.
%For the case of 1 qubit, the existence of PCE generators $\pceg$ is  
%verified by Siudzińska~\cite{siudzinska2019geometry}.
%\janote{Esto puede
%ser un párrafo, o puede ser un comentario final dentro del párrafo antrerior.}

\cpnote{Arreglar}
PCE generators may be identified with a unique vector index
$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
the identity map and the 3 1-qubit PCE quantum channels that project the 
Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
Therefore we will correspond every PCE generator to a collection of numbers 
that tells what its local-action is on every individual qubit in the system.
All of this collections of numbers happen to be all vector indices $\{\valpha\}$.
For example, the PCE generator represented in \Fref{fig:examplesPCE}(c),
whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
to the $x$ axis and on the second qubit (first row) to map the corresponding
local Bloch sphere to itself, may be characterized with the vector index (1,0). 
Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
every PCE generator and so denote them as $\pceg$.

\cpnote{Mucha lora y no se prueba nada}
PCE generators $\pceg$ obey a symmetry
that defines uniquely every PCE generator. The coefficients  $\{\taus\}$ of every 
generator $\pceg$ arranged in a hypercubic $N$-dimensional array 
%of dimensions $\underbrace{4\times 4\times\ldots\times 4}_{N}$ 
obey a symmetry with respect to the $N$ hyperplanes of dimension $N-1$
\janote{según wikipedia, por definición, los hiperplanos son de dimensión 
$N-1$. Así que podría quitarse la aclaración de la dimensión. Lo dejo así y 
espero retroalimentación} 
that separate the array into two half arrays \janote{No estoy tan satisfecho con la redacción de este enunciado}.
See, in \Fref{fig:examplesPCE}, that generators $\mcG_{(1,3)}$, $\mcG_{(3,2)}$
and $\mcG_{(2,2)}$ represented in (c), (d) and (e), respectively,
are either symmetric or anti-symmetric with respect to the lines that cut vertically 
and horizontally in half the diagrams. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows.\janote{sintetizar} Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.

>>>>

Un párrafo de intro a las tres ideas 
\begin{itemize}
\item Existen los generadores 
\item Clasificación de los generadores 
\item Simetrías
\end{itemize}

In this section we will discuss the existence of a generator set for all PCE
quantum channels, how to label each element of this set according to its local-action to 
every qubit in the system, and, finally, we  	will discuss a symmetry of
both PCE quantum channel generators and $A$, the matrix that diagonalizes
the Choi matrix $\mcD_N$ of a PCE map.

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs can be obtained as
compositions of PCEs corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCEs that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
% and the complete vector space (corresponding to the identity channel),
% which we call generators. 
% \cpnote{Aca tenemos un semigrupo y no se si incluir a la identidad como 
% generador. De alguna manera es excepcional. Es problematico incluirla y es 
% problematico no incluirla.}
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
PCE generators and the composition of any two of them yield the PCE channel
corresponding to (b). 
% 
% 
% the composition of any two 2-qubits PCE
% generators 
% in (c), (d), or (e) is equivalent to taking the intersection between them both
% and the three different possible compositions result in PCE channel in (b).
%\cpnote{Aca se me hace que hay un cambio brusco de tema. Ver que onda
%con Jose Alfredo}
%The existence of the PCE generators $\mcG$ provide the program of 
%constructing all PCE channels a tool that avoids the usage of a 
%high-computational-cost method. In a moment, we will discuss how to find the 
%generators $\mcG$ from matrix $A$, given a symmetry that is obeyed both by the generators 
%$\mcG$ and matrix $A$ in \Eref{eq:san}\cpnote{Como que para que ahora se
%menciona entonces?}.
%For the case of 1 qubit, the existence of PCE generators $\pceg$ is  
%verified by Siudzińska~\cite{siudzinska2019geometry}.
%\janote{Esto puede
%ser un párrafo, o puede ser un comentario final dentro del párrafo antrerior.}

\cpnote{Arreglar}
PCE generators may be characterized by its local-action on every qubit 
in the system, and thus be denoted as $\mcG_{\vec\alpha}$. First, let us denote
the four 1-qubit PCE generators as $\mcG_{\alpha}$, with $\alpha=0,1,2,3$.
The first being the trivial generator and the remaining the channels that 
map the Bloch sphere to each of the Cartesian axes.
%PCE generator may be identified with a unique vector index
%$\vec\alpha$, and thus be denoted as $\pceg$.
% El numero de generadores concide con esa vaina. 
%Let us introduce a convention for the following discussion for 
%the indices in $\vec\alpha$; the number 0 to 3 identifies
%Recall that indices $0,1,2,3$ in $\vec\alpha$ correspond to, respectively,
%the identity map and the 3 1-qubit PCE quantum channels that project the 
%Bloch sphere to the rectangular axes $x$, $y$ and $z$. 
%If one sets all indices in $\vec\alpha$ to zero except for the $k$-th index.
\janote{A rigorous explanation of why every $\vec\alpha$ is corresponded
to every generator.}
Therefore we can identify the index $\alpha_k$ in $\vec\alpha$ with the local-action
of the PCE generator on the $k$-th qubit. In consequence, we will denote 
PCE generators as $\mcG_{\vec\alpha}$. 
See in \Fref{fig:appex:twoQ_PCEGs} the notation of  two-qubit PCE generators.
%See, for example, non-trivial generators in subfigures (c), (d) and (e) of
%\Fref{fig:examplesPCE} are denoted $\mcG_{(1,3)}$, $\mcG_{(3,2)}$ and
%$\mcG_{(2,2)}$, respectively.
%whose local-action on the first qubit (first column) is to project the corresponding local Bloch sphere 
%to the $x$ axis and on the second qubit (first row) to map the corresponding
%local Bloch sphere to itself, may be characterized with the vector index (1,0). 
%Then, it is clear that in general one can associate a unique $\vec\alpha$ to 
%every PCE generator and so denote them as $\pceg$.

\cpnote{Mucha lora y no se prueba nada}
The same symmetry can be identified for both the PCE generators $\pceg$ and
matrix $A$ the diagonalizes PCE map's Choi matrix $\mcD_N$.
%PCE generators $\pceg$ obey a symmetry
%that defines uniquely every PCE generator. 
For a PCE generator the symmetry is encoded in the array of $\taus$, and 
for $A$ in its rows and columns.\janote{aquí voy}
The coefficients  $\{\taus\}$ of every 
generator $\pceg$ arranged in a hypercubic $N$-dimensional array 
%of dimensions $\underbrace{4\times 4\times\ldots\times 4}_{N}$ 
obey a symmetry with respect to the $N$ hyperplanes of dimension $N-1$
\janote{según wikipedia, por definición, los hiperplanos son de dimensión 
$N-1$. Así que podría quitarse la aclaración de la dimensión. Lo dejo así y 
espero retroalimentación} 
that separate the array into two half arrays \janote{No estoy tan satisfecho con la redacción de este enunciado}.
See, in \Fref{fig:examplesPCE}, that generators $\mcG_{(1,3)}$, $\mcG_{(3,2)}$
and $\mcG_{(2,2)}$ represented in (c), (d) and (e), respectively,
are either symmetric or anti-symmetric with respect to the lines that cut vertically 
and horizontally in half the diagrams. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows.\janote{sintetizar} Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.
